---
title: "Analysis of NME rates in the United States"
output: html_notebook
---

Note that the analysis only is performed in this notebook. 
Figures 1 and 2 were produced using Python, with the same dataset (figure 1) or with a dataset using values predicted by the analysis in this notebook (figure 2). 

First import the required packages used throughout the notebook. Make sure these are installed using `install.packages()`.

```{r warning = FALSE, message = FALSE}
rm(list = ls())
library(conflicted)
library(tidyverse)
library(boot)
library(Kendall)
library(betareg)
library(RCurl)
```

##Analysis of the association of NME proportions and vaccination rates in the United States in school year 2016-2017 and 2017-2018
This corresponds to manuscript Figure 1 and supplemental figure 1

Import the dataset

```{r}
#Full dataset - includes both school years
#Note that this is linked to the Github repository and does not required the data to be downloaded. 
#This would need to be changed if the data is imported from a local source.
data_fig1 <- read_csv(getURL('https://raw.githubusercontent.com/Rom1Garnier/NME/master/Data/data_fig1_final.csv'))
```

```{r warning = FALSE}
#List of vaccines and years - to loop on
Vaccines <- c('MMR', 'DTaP', 'Varicella')
Years <- c(2016, 2017)

#Create an empty dataframe to put the results in
result_table <- data.frame('Year' = numeric(), 'Vaccine' = character(), 'Beta_regression_estimate' = numeric(), 'Beta_regression_pvalue' = numeric())

#Loop over vaccines and years, and run the Spearman correlation and the Beta regression for each year/vaccine pair
for(vac in Vaccines){
  for(year in Years){
    data_temp <- data_fig1 %>% 
      dplyr::filter(Year == year, Vaccine == vac)
      Betareg_mod <- betareg(Vaccination_rate ~ p_NME_tot, data = data_temp)
    
    #Create a dataframe with the result of the current year/vaccine pair
    result_models <- data.frame('Year' = year, 'Vaccine' = vac, 'Beta_regression_estimate' = round(Betareg_mod$coefficients$mean[2], 3), 'Beta_regression_pvalue' = round(summary(Betareg_mod)$coefficients$mean[8], 3))
    
    #Place the results in the result table
    result_table <- result_table %>% bind_rows(result_models)
  }
}
  
```

##Dynamics of ME and NME rates in select states with a history of recent policy changes 

Import the data first 

```{r}
#Note that this points to the location of the data in the Github repository associated with the study
data_ts_NME <- read_csv(getURL('https://raw.githubusercontent.com/Rom1Garnier/NME/master/Data/data_fig2.csv'))
```

Now run the forecasted linear models for years before the policy change 

```{r}
#Create empty dataframe
pred <- as.data.frame(matrix(ncol = 0, nrow = 0))

#Loop over the states in the dataframe
for(state in unique(data_ts_NME$State)){
  
  #Exctract the data for a given state
  data_temp <- data_ts_NME %>% 
    dplyr::filter(State == state)
  
  #Restrict the data to the years prior to the policy change
  data_mod <- data_temp %>% 
    dplyr::filter(Year_begin < Year_change)
  
  #Run two models - for personal belief exemptions (PBE), and medical exemptions (PME)
  mod_PBE <- lm(Percent_PBE ~ Year_begin, data = data_mod)
  mod_PME <- lm(Percent_PME ~ Year_begin, data = data_mod)
  
  #Place the predicted values in the dataframe
  data_temp$Mod_pred_PBE <- stats::predict(mod_PBE, data.frame(Year_begin = data_temp$Year_begin))
  data_temp$Mod_pred_PME <- stats::predict(mod_PME, data.frame(Year_begin = data_temp$Year_begin))
  
  #In the case of Vermont also run independent models for religious and philosophical exemptions
  if(state == 'Vermont'){ 
    mod_philo <- lm(Percent_PBE_philo ~ Year_begin, data = data_mod)
    mod_rel <- lm(Percent_PBE_religious ~ Year_begin, data = data_mod)
  
    data_temp$Mod_pred_Philo <- stats::predict(mod_philo, data.frame(Year_begin = data_temp$Year_begin))
    data_temp$Mod_pred_Religious <- stats::predict(mod_rel, data.frame(Year_begin = data_temp$Year_begin))
    }

  pred <- pred %>% 
    bind_rows(data_temp)
}

folder_path <- '' #provide a local path to export the dataset with predicted values. 
write_csv(pred, paste(folder_path, 'data_fig2_with_predictions.csv'))
```

